---
title: "Tables for Personal Finance"
author: "Deepansh Khurana"
format:
  html:
    code-fold: true
theme:
  - cosmo
  - style/custom.scss
editor: visual
---

::: {style="text-align: justify;"}
When I made the [Hrafnagud](https://github.com/DeepanshKhurana/Hrafnagud-Dynamo "Hrafnagud-Dynamo") project I recently talked about at ShinyConf (and am about to talk about at useR! in July), I also made several apps for myself to consume each API. The app and API for tracking my finances is called **Ebenezer**, named after the miserly Scrooge. My goal was to take each kind of financial instrument and make a table that fits well for it. Bonds work differently than Equity stocks, for example, and therefore, should be represented differently. The result was the following tables, which are now a part and parcel of my application.
:::

#### Key Features

::: {style="text-align: justify;"}
One of the key features of these tables are the way the colors are assigned automatically. This is done using a *get_css_class()* function which basically assigns the different colors and highlights wherever required.

Another important bit is how the tables themselves are constructed. Processing functions are written in the *cell* parameter of the *reactable()* call. The function then takes up a concatenated string and builds up a *div* structure that is styled with a custom scss file.
:::

## Setup

Here we setup some libraries, functions and load the dataframes for these examples.

```{r "setup"}

suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(reactable))
suppressPackageStartupMessages(library(reactablefmtr))
suppressPackageStartupMessages(library(shiny))

source("utils/data_formatting.R")

files <- list.files("data", "*.csv")
paths <- paste("data", files, sep = "/")
data <- lapply(
  paths,
  read.csv
)

names(data) <- gsub(".csv", "", files)
```

Now that the data is loaded, we can take each table at a time.

### Ticker

::: {style="text-align: justify;"}
The ticker is a simple table that needs a few key pieces of information. The most important out of which is the correct semantic colour for rise and fall of a stock.

**Note:** While I usually consume this data from a custom API, for the sake of illustration and reproducibility, I have included a static file here as a .csv.

Let's try to make a Ticker Table.
:::

```{r "ticker"}

ticker_data <- data$ticker_data

colnames(ticker_data) <- c(
  "Code", "Stock", "Price", "Change", "Percent"
)

ticker_data$Stock <- paste(
  ticker_data$Code,
  ticker_data$Stock,
  sep = "_"
)

ticker_data$Price <-
  paste(ticker_data$Price,
    ticker_data$Change,
    ticker_data$Percent,
    sep = "_"
  )

ticker_data <- ticker_data[, c("Stock", "Price")] |>
  na.omit()

reactable(
  data = ticker_data,
  pagination = FALSE,
  sortable = FALSE,
  searchable = TRUE,
  columns = list(
    Stock = colDef(
      align = "left",
      cell = function(value) {
        value <- strsplit(value, "_")[[1]]
        div(
          class = "stock-name-row",
          div(
            class = "stock-code",
            value[1]
          ),
          div(
            class = "stock-name",
            value[2]
          )
        )
      }
    ),
    Price = colDef(
      align = "right",
      cell = function(value) {
        value <- strsplit(value, "_")[[1]]

        price <- as.numeric(value[1])
        change <- as.numeric(value[2])
        percent <- as.numeric(value[3])

        indicator <- get_css_class(change, type = "all")

        div(
          class = "stock-price-row",
          div(
            class = "stock-price",
            style = list(
              fontSize = "1.5em",
              fontWeight = 600
            ),
            paste0(
              "$",
              sub(
                "\\.0+$",
                "",
                format_price(price)
              )
            )
          ),
          div(
            class = "stock-change-block",
            div(
              class = paste("stock-percent",
                indicator$text,
                indicator$highlight,
                sep = " "
              ),
              paste0(sub("-", "", preserve_zero(percent)), "%")
            ),
            div(
              class = paste("stock-change",
                indicator$text,
                sep = " "
              ),
              sub("-", "", format_price(change))
            )
          ),
        )
      }
    )
  )
)
```

### Stocks

::: {style="text-align: justify;"}
Alright, now that we have data for the ticker, we can create a stocks table. The table below gives you the portfolio with different fields such as Invested Value, Quantity, Average Price on the left, and Unrealized P&L, Realized P&L, Current Price on the right. If a user hovers on the icons, they will find these labels as well.
:::

```{r "stocks"}

portfolio <- data$stocks_data |>
        mutate(
          unrealized_percent =
            true_round(
              unrealized * 100 / holding_value, 2
            )
        ) |>
        mutate(
          Left =
            paste(
              stock_symbol,
              name,
              holding_value,
              quantity,
              avg_price,
              sep = "_"
            ),
          Right = paste(
            unrealized,
            unrealized_percent,
            realized,
            current_price,
            change_percent,
            sep = "_"
          )
        ) |>
        select(Left, Right)

reactable(
  data = portfolio,
  pagination = FALSE,
  searchable = TRUE,
  columns = list(
    Left = colDef(
      align = "left",
      cell = function(value) {
        value <- strsplit(value, "_")[[1]]
        div(
          class = "stock-left-area",
          div(
            class = "stock-name-row",
            div(
              class = "stock-code",
              value[1]
            ),
            div(
              class = "stock-name",
              value[2]
            )
          ),
          div(
            class = "stock-holding-row",
            div(
              class = "stock-market-value small-row-left",
              title = "Invested Value",
              icon(name = "vault", class = "label"),
              p(class = "value", format_price(value[3], 2))
            ),
            div(
              class = "stock-quantity small-row-left",
              title = "Quantity",
              icon(name = "hashtag", class = "label"),
              p(class = "value", value[4])
            ),
            div(
              class = "stock-avg-price small-row-left",
              title = "Average Price",
              icon(name = "barcode", class = "label"),
              p(class = "value", paste0("₹", true_round(value[5], 2)))
            )
          )
        )
      }
    ),
    Right = colDef(
      align = "right",
      cell = function(value) {
        value <- strsplit(value, "_")[[1]]

        unrealized_class <- get_css_class(value[2], type = "text")
        realized_class <- get_css_class(value[3], type = "text")
        current_price_class <- get_css_class(value[5], type = "text")

        div(
          class = "stock-right-area",
          div(
            class = "stock-unrealized-row small-row-right",
            title = "Unrealized P&L",
            icon(
              name = "money-bill-transfer",
              class = paste("stock-current-price label",
                unrealized_class,
                sep = " "
              )
            ),
            p(
              class = paste("stock-unrealized value",
                unrealized_class,
                sep = " "
              ),
              format_price(value[1])
            ),
            p(
              class = paste("stock-unrealized-percent value",
                unrealized_class,
                sep = " "
              ),
              paste0("(", value[2], "%)")
            )
          ),
          div(
            class = "stock-realized small-row-right",
            title = "Realized P&L",
            icon(
              name = "receipt",
              class = paste("stock-current-price label",
                realized_class,
                sep = " "
              )
            ),
            p(
              class = paste("value",
                realized_class,
                sep = " "
              ),
              format_price(value[3])
            )
          ),
          div(
            class = "stock-current-price-row small-row-right",
            title = "Current Price",
            icon(
              name = "comment-dollar",
              class = paste("stock-current-price label",
                current_price_class,
                sep = " "
              )
            ),
            div(
              p(
                class = paste(
                  "stock-current-price value",
                  current_price_class,
                  sep = " "
                ),
                value[4]
              )
            ),
            p(
              class = paste(
                "stock-current-price-percent value",
                current_price_class,
                sep = " "
              ),
              paste0("(", value[5], "%)")
            )
          )
        )
      }
    )
  )
)
```

### Bonds

::: {style="text-align: justify;"}
Between Stocks and Bonds, something changes. Most stock data is current. You need the most crucial information immediately. With bonds, the information needed is a bit different. You need to know the current amount, the amount it will total to when the time period passes and the progress. So, with the help of *reactablefmtr*, I added a progress bar to the tables.

Another thing to note is that the progress bar color changes according to a palette. This gives the user visual feedback for how far their bonds have progressed to maturity. Darker is "matures soon".
:::

```{r "bonds"}

bonds <- data$sgb_data

today <- Sys.Date()

bonds$info <- paste(
  bonds$name,
  bonds$invested_date,
  bonds$maturity_date,
  bonds$category,
  bonds$quantity,
  bonds$unit_price,
  bonds$total,
  sep = "_"
)

bonds$progress <-
  1 - as.numeric(
    difftime(
      format_date(bonds$maturity_date),
      as.Date(today),
      units = "days"
    )
  ) / (8 * 365.25)

bonds <- bonds |>
  select(info, progress) |>
  rowwise() |>
  mutate(
    color = get_progress_bar_colour(
      progress,
      c(
        "#a3b18a",
        "#588157",
        "#3a5a40",
        "#344e41"
      )
    )
  ) |>
  ungroup()

reactable(
  bonds,
  class = "sgb-reactable",
  columns = list(
    info = colDef(
      name = "Bond",
      cell = function(string) {
        string <- strsplit(string, "_")[[1]]
        div(
          class = "sgb-row",
          div(
            class = "sgb-name",
            gsub(
              "BOND NAME",
              "/",
              string[1]
            )
          ),
          div(
            class = "sgb-timeline",
            year(format_date(string[2])),
            icon("arrow-right"),
            year(format_date(string[3]))
          ),
          div(
            class = "sgb-metrics-container",
            div(
              class = "sgb-category metric",
              title = "Category",
              icon("receipt", class = "label"),
              p(string[4], class = "value")
            ),
            div(
              class = "sgb-quantity metric",
              title = "Quantity",
              icon("hashtag", class = "label"),
              p(string[5], class = "value")
            ),
            div(
              class = "sgb-price metric",
              title = "Unit Price",
              icon("barcode", class = "label"),
              p(format_price(string[6]), class = "value")
            ),
            div(
              class = "sgb-total metric",
              title = "Total Invested",
              icon("vault", class = "label"),
              p(format_price(string[7]), class = "value")
            ),
          )
        )
      }
    ),
    progress = colDef(
      cell = data_bars(
        bonds,
        max_value = 1,
        fill_color_ref = "color",
        background = "#f1f3f4",
        round_edges = TRUE,
        bar_height = 20,
        number_fmt = scales::label_percent()
      )
    ),
    color = colDef(
      show = FALSE
    )
  )
)
```
